---
title: "census_data_load_weighting"
output: html_document
date: '2022-06-20'
---

# 2016 ACS 
```{r}
variables_to_get <- c(
  population = "B01003_001",
  median_income = "B19013_001",
  zero_vehicle_count = "B08201_002", 
  total_hh_vehicle = "B08201_001",
  owner_occ_hh = "B08137_002",
  total_hh_occ = "B08137_001", 
  med_gross_rent = "B25031_001", 
  median_age = "B01002_001",
  bachelors = "B15003_022",
  white = "B02001_002",
  black = "B02001_003",
  native = "B02001_004",
  asian = "B02001_005",
  pi = "B02001_006",
  latino = "B03003_003"
)

# 2016 5yrs ACS
nyc_acs <- get_acs(
  geography = "tract",
  variables = variables_to_get,
  state = "NY",
  county = "New York",
  survey = "acs5",
  key = , # need to add your own
  geometry = TRUE,
  output = "wide",
  year = 2016
)
```
```{r}
# get rid of margin of error columns
nyc_acs = nyc_acs %>% select(!ends_with("M"))
```

# Read in taxi zones for population interpolation
```{r}
# read in TZs
tz_shp <- readOGR(dsn = "../documents/taxi_zones", layer = "taxi_zones")
tz_shp = subset(tz_shp, borough == "Manhattan")
tz_shp <- spTransform(tz_shp, CRS("+init=epsg:4269 +proj=longlat +ellps=GRS80 +datum=NAD83
+no_defs +towgs84=0,0,0"))
tz_shp = st_as_sf(tz_shp) # convert to special features 
```

# use census blocks - more granular than tracts - for population aggregation to allocate influence to TZs
```{r}
st_crs(tz_shp)
st_crs(nyc_acs)

nyc_blocks <- blocks(
  state = "NY",
  county = "New York",
  year = 2016
)

# 2016 5yrs ACS
pop_2016_acs <- get_acs(
  geography = "block group",
  variables = c(population = "B01003_001"),
  state = "NY",
  county = "New York",
  survey = "acs5",
  key = "",
  geometry = TRUE,
  output = "wide",
  year = 2016
)

st_point_on_surface(pop_2016_acs)

class(pop_2016_acs$populationE)

pop_2016_acs = pop_2016_acs[!st_is_empty(pop_2016_acs),,drop=FALSE]
pop_2016_acs = as.data.frame(pop_2016_acs)
pop_2016_acs = pop_2016_acs %>% select(GEOID, populationE, geometry)

colnames(pop_2016_acs)[which(names(pop_2016_acs) == "populationE")] <- "pop16"

pop_2016_acs = st_as_sf(pop_2016_acs)
```

# population interpolation demographics
```{r}
sums_df = nyc_acs %>% select(-c(median_incomeE, med_gross_rentE, median_ageE))

# sums
interp_sums = interpolate_pw(
  sums_df,
  tz_shp,
  to_id = "LocationID",
  extensive = TRUE, 
  weights = pop_2016_acs,
  weight_column = "pop16"
)

averages_df = nyc_acs %>% select(median_incomeE, med_gross_rentE, median_ageE)

# averages
interp_avgs = interpolate_pw(
  averages_df,
  tz_shp,
  to_id = "LocationID",
  extensive = F, 
  weights = pop_2016_acs,
  weight_column = "pop16"
)
```


# Merge the two sf dataframes, one with sum interpolation and one with avgs
```{r}
# convert to dfs to get rid of geometry in one of the objects and merge based on locationid
interp_avgs_df = as.data.frame(interp_avgs)
interp_avgs_df = interp_avgs_df %>% select(-geometry)

interp_sums_df = as.data.frame(interp_sums)

pop_interp = merge(interp_sums_df, interp_avgs_df, by = "LocationID")
pop_interp = st_as_sf(pop_interp)
```

# create pct owner occ, pct college, pop_density AFTER getting sums based on population interpolated to the TZ level
```{r}
pop_interp = pop_interp %>% mutate(pct_zero_vehicle = zero_vehicle_countE/total_hh_vehicleE,
                   pct_owner_occ = owner_occ_hhE/total_hh_occE,
                   pct_bachelors = (bachelorsE) / populationE,
                   pop_density = as.numeric(set_units(populationE / st_area(.), "1/km2")),
                   white_pct = whiteE/populationE,
                   black_pct = blackE/populationE,
                   native_pct = nativeE/populationE,
                   asian_pct = asianE/populationE,
                   latino_pct = latinoE/populationE)
```

# merge census data with rides, and get rid of census counts bc we created percentages (e.g., pct zero vehicle) above
```{r}
census_rides <- merge(full_data, pop_interp, by.x = "locationID", by.y = "LocationID")

census_rides %>% filter(locationID %in% c(75, 262, 263, 140, 141, 236, 237)) %>% arrange(month)

census_rides = census_rides %>% select(-c(zero_vehicle_countE, total_hh_vehicleE, owner_occ_hhE, total_hh_occE, bachelorsE, whiteE, blackE, nativeE, asianE, piE, latinoE))
```

