---
title: "regression_model"
output: html_document
date: '2022-06-20'
---

```{r}
library(spdep)
library(plm)
library(lmtest)
library(splm)
library(finalfit)
library(dplyr)
library(jtools) 
```
# ensure model data is in correct order for panel regression
```{r}
model_data = model_data[order(model_data$locationID, model_data$month), ]
```

# OLS model
```{r}
OLS_formula <- "observations ~ populationE + pop_density + median_incomeE + med_gross_rentE + median_ageE + pct_bachelors + pct_zero_vehicle + pct_owner_occ + black_pct + latino_pct + asian_pct + pct_com + pct_other + Treatment:pre_post"
OLS_model <- lm(formula = OLS_formula, data = model_data, na.action=na.exclude)
summary(OLS_model)
```

# look for colinearity - over 5 indicates problematic 
```{r}
vif(OLS_model)
```


```{r}
OLS_formula2 <- "observations ~ populationE + pop_density + median_incomeE + median_ageE + pct_bachelors + pct_zero_vehicle + pct_owner_occ + black_pct + latino_pct + asian_pct + pct_com + pct_other + Treatment:pre_post"
OLS_model2 <- lm(formula = OLS_formula2, data = model_data, na.action=na.exclude)
summary(OLS_model2)
```

# check colineairty 
```{r}
vif(OLS_model2)

# correlations 
cor(model_data$pct_bachelors, model_data$median_incomeE)

cor(model_data$populationE, model_data$pop_density)

cor(model_data$median_incomeE, model_data$pct_owner_occ) # remove pct_owner_occ becuase of colinearity with income 

cor(model_data$latino_pct, model_data$black_pct) # KEEP even though high correlation becuase
```

# Final OLS model 

```{r}
OLS_formula3 <- "observations ~ populationE + median_incomeE + median_ageE + pct_zero_vehicle + black_pct + latino_pct + asian_pct + pct_com + pct_other + Treatment:pre_post"
OLS_model3 <- lm(formula = OLS_formula3, data = model_data, na.action=na.exclude)
summary(OLS_model3)
vif(OLS_model3)
```

# Fixed effects approach #1: Add a dummy var for each T and each TZ by using factor. 
```{r}
dummy_formula <- "observations ~ populationE + median_incomeE + median_ageE + pct_zero_vehicle + black_pct + latino_pct + asian_pct + pct_com + pct_other + Treatment:pre_post + factor(month) + factor(locationID)-1"
dummy_model <- lm(formula = dummy_formula, data = model_data, na.action = na.omit)
summary(dummy_model)
```

# Fixed effects approach #2: within estimator

gets rid of the terms for each state, and therefore intercepts for each state. Uses plm, or panel linear model.

```{r}
# WE_formula <- "observations ~ populationE + median_incomeE + median_ageE + pct_bachelors + pct_zero_vehicle + pct_owner_occ + black_pct + pct_res + pct_other + Treatment:pre_post"

# two-way
# WE_model = plm(WE_formula,
#                data = model_data,
#                index = c("locationID", "month"),
#                model = "within",
#                effect = "twoways")
# summary(WE_model)


WE_formula <- "observations ~ populationE + median_incomeE + median_ageE + pct_bachelors + pct_zero_vehicle + pct_owner_occ + black_pct + pct_res + pct_other + Treatment:pre_post + factor(month)"

WE_model = plm(WE_formula,
               data = model_data,
               index = "locationID",
               model = "within")
summary(WE_model)
```

# F test to check whether the presence of fixed effects improves the model. ie, F test for additional effects on top of OLS
```{r}
pFtest(WE_model, OLS_model) 
```

# check for normal distribution of errors 
```{r}
p = ggplot(as.data.frame(residuals(WE_model)), aes(x = residuals(WE_model))) + 
  geom_histogram(bins = 100, alpha = 0.5, color = "navy",
                 fill = "navy") + xlab("Error") + ylab("Count")
p = p + ggtitle("Within-Estimator Model Residuals") + 
                    theme(plot.title = element_text("Helvetica", "bold", "black", "12", vjust = 2)) + theme_minimal()
p
```

Two assumptions:
Normal distribution - good. Errors are not biased towards one side. 
Independence of observations - let's check. 

# CD and RW tests
```{r}
model_data$residuals <- residuals(WE_model)

pcdtest((pdata.frame(model_data))$observations, index = c("month"), w = tz_mat)
pcdtest((pdata.frame(model_data))$residuals, index = c("month"), w = tz_mat)

rwtest((pdata.frame(model_data))$observations, w = tz_mat, replications = 999)
rwtest((pdata.frame(model_data))$residuals, w = tz_mat, replications = 999)
```

# SE models
```{r}
# with all of the terms included, I get the error: Error in solve.default(t(X) %*% sigma.1 %*% X, t(X) %*% sigma.1 %*% y) : system is computationally singular. This indicates a problem with multicollinearity. 

SEM_data <- pdata.frame(model_data, index = c("locationID")) 

SEM_fe_formula = "observations ~ locationID + factor(month) + Treatment:pre_post"
  
SEM_fe <- spreml(SEM_fe_formula, data = SEM_data, w = tz_listw,
lag = FALSE, errors = "sem")
summary(SEM_fe, digits = 2)

# With neighborhood fixed effects, I cannot add any varaible besides the DiD term. So this is a weaker model than the one that only controls for time FEs, whcih we know are significant. Forget this model. 

SEM_tz_fe_formula = "observations ~ locationID + Treatment:pre_post"

SEM_tz_fe <- spreml(SEM_tz_fe_formula, data = SEM_data, w = tz_listw,
lag = FALSE, errors = "sem")
summary(SEM_tz_fe)

# With time fixed effects, I can add all of the EVs.

SEM_month_fe_formula = "observations ~ factor(month) + populationE + median_incomeE + median_ageE + pct_bachelors + pct_zero_vehicle + pct_owner_occ + black_pct + pct_com + pct_other + Treatment:pre_post"

SEM_month_fe <- spreml(SEM_month_fe_formula, data = SEM_data, w = tz_listw,
lag = FALSE, errors = "sem")
summary(SEM_month_fe)
```

# SEM model 3 residuals 
```{r}
p = ggplot(as.data.frame(residuals(SEM_month_fe)), aes(x = residuals(SEM_month_fe))) + 
  geom_histogram(bins = 100, alpha = 0.5, color = "navy",
                 fill = "navy") + xlab("Error") + ylab("Count")
p = p + ggtitle("Spatial Error Model (With Time Fixed Effects) Residuals") + 
                    theme(plot.title = element_text("Helvetica", "bold", "black", "12", vjust = 2)) + theme_minimal()
p
```
